# Bayesian Network Learning
## On the Default and Titanic data sets
### Karen Mazidi

### Default data set

Let's try bnlearn on some familiar data sets. First, the Default data set from ISLR. 

```{r}
library(bnlearn)
library(ISLR)
data("Default")
```

### Build the model

Use the hill-climbing algorithm to create the model, then display it. 

The model seems to fit what we learned by doing ML on this data set before. Income is not a good predictor for default, notice there is no link between them.


```{r}
bn_df <- data.frame(Default)
str(bn_df)
res <- hc(bn_df)
plot(res)
```


### Probabilities

This is a really unbalanced data set!

```{r}
fittedbn <- bn.fit(res, data=bn_df)
print(fittedbn$default)
```


### Try some predictions

```{r}
cpquery(fittedbn, event=(default=="Yes"), evidence = (student=="Yes"))
cpquery(fittedbn, event=(default=="Yes"), evidence = (balance>1200))
cpquery(fittedbn, event=(default=="Yes"), evidence = (student=="Yes" & balance>1200))
```

### Titanic

Now let's try a subset of the Titanic data set. We had to a little data fiddling because the bnlearn didn't like NAs and ints. 


```{r}
df <- read.csv('train.csv', header=T, na.strings=c(""))
df <- df[1:890, c(2,3,5,6)]  # Survived, Pclass, Sex, Age
bn_titan <- df[complete.cases(df), ]
bn_titan$Survived <- as.factor(bn_titan$Survived)
bn_titan$Pclass <- as.factor(bn_titan$Pclass)
bn_titan$Sex <- as.factor(bn_titan$Sex)
str(bn_titan)
```


### Build the net

This model is not surprising. Age and Sex and Pclass were found to be good predictors in previous experiments.

```{r}
res_titan <- hc(bn_titan)
plot(res_titan)
```

### Print the probabilities

This breaks down the probabilities nicely.


```{r}
fitted_bn_titan <- bn.fit(res_titan, data=bn_titan)
print(fitted_bn_titan)
```


### Predict

Try some predictions based on the net

```{r}
cpquery(fitted_bn_titan, event = (Survived==1), evidence = (Pclass==1))
cpquery(fitted_bn_titan, event = (Survived==1), evidence = (Age<9) )
cpquery(fitted_bn_titan, event = (Survived==1), evidence = (Pclass==1 & Age<=9) )
cpquery(fitted_bn_titan, event = (Survived==1), evidence = (Pclass==1 & Age>9) )
cpquery(fitted_bn_titan, event = (Survived==1), evidence = (Sex=="female") )
cpquery(fitted_bn_titan, event = (Survived==1), evidence = (Sex=="male" & Age>21) )
```


