---
title: "Using R Control Structures"
author: "Karen Mazidi"
output:
  pdf_document: default
  html_document: default
---

# Exploring R Control Structures with the PimaIndiansDiabetes2 data

First we load the mlbench package and then the diabetes data set. The data has 768 observations of 9 variables, which we can see with the structure str() function.  The data was collected from Pima Indian women in the late 1990s. 

```{r}
library(mlbench)
data(PimaIndiansDiabetes2)
df <- PimaIndiansDiabetes2   # df points to the data frame
str(df)
```

### Look for NAs

The sapply() function applies a function to elements of a list. In this case the elements of the list are columns in our data frame. The function is an anonymous function (we didn't name it), and it just sums, ignoring NAs.

```{r}
sapply(PimaIndiansDiabetes2, function(x) sum(is.na(x)))
```

### Write a function

As an example of how to write a function, we write a function named fill_NA that takes two arguments and returns a vector. In R the return() statement is often not needed since R will return the last thing evaluated. In this function, the mean_med variable is a switch to choose whether to fill NAs with the mean or the median. 

After the function is defined, we can call it with different columns. Then we use the complete.cases() function to get rid of the remaining rows that have NAs.

```{r}
fill_NA <- function(mean_med, v){
  # fill missing values with either 1=mean or 2=median
  if (mean_med == 1){
    m <- mean(v, na.rm=TRUE)
  } else {
    m <- median(v, na.rm=TRUE)
  }
  v[is.na(v)] <- m
  v
}

# make a new data set with NA's filled
df <- PimaIndiansDiabetes2[]
df$triceps <- fill_NA(1, df$triceps)
df$insulin <- fill_NA(1, df$insulin)
df <- df[complete.cases(df),]

```





### Plots and for loops

The following plots some data from the data set and 3 ablines which are 3 linear regression lines created by lm(). Each regression line is a different color and we added a legend.

```{r}
cols <- c(6,8,1)
plot(df$mass, df$glucose, main="PimaIndianDiabetes2")
for (i in 1:3){
    model <- lm(glucose~df[,cols[i]], data=df)[1]
    abline(model, col=i)
}
legend("topright", title="Predictors", c("Mass", "Age", "Pregnant"), fill=c(1,2,3))
```

### Using ifelse()

We use ifelse() to set a variable=1 if insulin is over 155 and 0 otherwise. Then we convert it to a factor, and place it in a new variable.

Then we plot. 

```{r}
df$large <- factor(ifelse(df$insulin>155,1,0))
plot(df$mass, df$glucose, pch=21, bg=c("blue","red")[unclass(df$large)])
```



