---
title: "Naive Baye from Scratch with the Titanic Data"
author: "Karen Mazidi"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---


We will use the same data and data cleaning as in the first notebook in this chapter, so we repeate those steps first with no commentary.

### Load the data

```{r}
df <- read.csv("data/titanic3.csv", header=TRUE, stringsAsFactors = FALSE)

# subset to just columns survived, pclass, sex, and age
df <- df[,c(1,2,4,5)]
# pclass and survived  and sex should be factors
df$pclass <- factor(df$pclass)
df$survived <- factor(df$survived)
df$sex <- factor(df$sex, levels=c("female", "male"))

# remove NAs
df <- df[!is.na(df$pclass),]
df <- df[!is.na(df$survived),]
df$age[is.na(df$age)] <- median(df$age,na.rm=T)

# divide into train and test
set.seed(1234)
i <- sample(1:nrow(df), 0.75*nrow(df), replace=FALSE)
train <- df[i,]
test <- df[-i,]

# perform Naive Bayes
library(e1071)
nb1 <- naiveBayes(df[,-2], df[,2], data=train)
pred <- predict(nb1, newdata=test[,-2], type="raw")

# look at first 5 (actual: 0 1 1 1 0)
pred[1:5,]
```



### Calculate priors

Using the training data we calculate prior probability fo survived the percentage for each category.

```{r}
apriori <- c(
  nrow(df[df$survived=="0",])/nrow(df),
  nrow(df[df$survived=="1",])/nrow(df)
  )
print("Prior probability, survived=no, survived=yes:")
apriori
```

### Calculate likelihoods for qualitative data

The likelihood for qualitative data is calculated as follows:

* for each class
* for each factor i
* likelihood (class=i|survived=yes) = count(factor = i and survived=yes) / count(survived=yes)
* likelihood (class=i|survived=no) = count(factor = i and survived=no) / count(survived=np)

we will use nrow() for our counts

```{r}
# get survived counts for no and yes
count_survived <- c(
  length(df$survived[df$survived=="0"]),
  length(df$survived[df$survived=="1"])
)
# likelihood for pclass
lh_pclass <- matrix(rep(0,6), ncol=3)
for (sv in c("0", "1")){
  for (pc in c("1","2","3")) {
    lh_pclass[as.integer(sv)+1, as.integer(pc)] <- 
      nrow(df[df$pclass==pc & df$survived==sv,]) / count_survived[as.integer(sv)+1]
  }
}

# likelihood for sex
lh_sex <- matrix(rep(0,4), ncol=2)
for (sv in c("0", "1")){
  for (sx in c(1, 2)) {
    lh_sex[as.integer(sv)+1, sx] <- 
      nrow(df[as.integer(df$sex)==sx & df$survived==sv,]) /
       count_survived[as.integer(sv)+1]
  }
}

```

### likelihood p(survived|pclass)

```{r}
print("Likelihood values for p(pclass|survived):")
lh_pclass
```

### likelihood p(survived|sex)

```{r}
print("Likelihood values for p(sex|survived):")
lh_sex
```


### Calculate likelihoods for quantitative data

Age is quantitative. We need to compute the mean and variance.

```{r}
age_mean <- c(0, 0)
age_var <- c(0, 0)
for (sv in c("0", "1")){
    age_mean[as.integer(sv)+1] <- 
      mean(df$age[df$survived==sv])
    age_var[as.integer(sv)+1] <- 
      var(df$age[df$survived==sv])
  }

```

### Probability density for quantitative data

For the qualitative variable we can calculate probailities by dividing but for the age variable we need a function that will calculate its probability.

```{r}
calc_age_lh <- function(v, mean_v, var_v){
  # run like this: calc_age_lh(6, 25.9, 138)
  1 / sqrt(2 * pi * var_v) * exp(-((v-mean_v)^2)/(2 * var_v))
}

```



### Function for scratch model

Write a function to calculate raw probabilities given pclass, sex, and age. 

```{r}
calc_raw_prob <- function(pclass, sex, age) {
  # pclass=1,2,3  sex=1,2   age=numeric
  num_s <- lh_pclass[2, pclass] * lh_sex[2, sex] * apriori[2] *
      calc_age_lh(age, age_mean[2], age_var[2])
  num_p <- lh_pclass[1, pclass] * lh_sex[1, sex] * apriori[1] *
      calc_age_lh(age, age_mean[1], age_var[1])
  denominator <- lh_pclass[2, pclass]  * lh_sex[2, sex] * calc_age_lh(age, age_mean[2], age_var[2]) * apriori[2] + 
      lh_pclass[1, pclass]  * lh_sex[1, sex] * calc_age_lh(age, age_mean[1], age_var[1]) * apriori[1]
  return (list(prob_survived <- num_s / denominator, prob_perished <- num_p / denominator))
}


```

### Apply to the first 5 test observations

Let's look at just the first 5 test observations. 


Note getting the right numbers, need to rethink likelihood values, flip table. Works for first 3=female, not for male.

```{r}
for (i in 1:5){
  raw <- calc_raw_prob(test[i,1], as.integer(test[i,3]), test[i,4])
  print(paste(raw[2], raw[1]))
}

pred[1:5,]
```

