---
output:
  pdf_document: default
editor_options: 
  chunk_output_type: console
---
# kNN Clustering - Regression
## Using 10-fold cross validations
### Karen Mazidi

Load the data

```{r}
library(ISLR)
df <- Auto[]
df$origin <- as.integer(df$origin)
# subset to columns mpg, weight, year, origin
df <- data.frame(scale(df[, c(1, 5, 7, 8)]  ))
```

### Create the 10 folds

We could do this manually but there is a function in caret that does this. Since the Auto data is a little less than 400 rows, we expect each of the 10 folds to be of legth 40 or less. We confirm that with sapply.


```{r}
library(caret)
set.seed(1234)
folds <- createFolds(df$mpg, k=10)
sapply(folds, length)
```

### Look at the fold indices

To get a better idea of the folds, let's just print the indices for each fold.

```{r}
for (i in 1:10){
  print(folds[[i]])
}
```

### Perform 10-fold cv

For now we will just let k=3 and perform 10-fold cv, then average the correlation and mse values.

```{r}
test_mse <- rep(0, 10)
test_cor <- rep(0, 10)
for (i in 1:10){
  fit <- knnreg(df[-folds[[i]], 2:4], df$mpg[-folds[[i]]], k=3)
  pred <- predict(fit, df[folds[[i]], 2:4])
  test_cor[i] <- cor(pred, df$mpg[folds[[i]]])
  test_mse[i] <- mean((pred - df$mpg[folds[[i]]])^2)
}
print(paste("Average correlation is ", round(mean(test_cor), 2)))
print(paste("range is ", range(test_cor)))
print(paste("Average mse is ", round(mean(test_mse), 2)))
print(paste("range is ", range(test_mse)))
```

### Try with various k

We modify the code above to be an anonymouse function called by sapply.  


```{r}
# try various values for k
k_values <- seq(1, 39, 2)
results <- sapply(k_values, function(k){
  mse_k <- rep(0, 10)
  cor_k <- rep(0, 10)
  for (i in 1:10){
    fit <- knnreg(df[-folds[[i]], 2:4], df$mpg[-folds[[i]]], k=k)
    pred <- predict(fit, df[folds[[i]], 2:4])
    cor_k[i] <- cor(pred, df$mpg[folds[[i]]])
    mse_k[i] <- mean((pred - df$mpg[folds[[i]]])^2)
  }
  #print(paste(mean(cor_k), mean(mse_k)))
  list(mean(cor_k), mean(mse_k))
})
# reshape results into matrix
m <- matrix(results, nrow=20, ncol=2, byrow=TRUE)
```


### Examine results

Plot the correlation and mse for each value of k.

```{r}
par(mfrow=c(2, 1))
plot(1:20, unlist(m[,1]), lwd=2, type="o", col='red', ylab="Correlation")
plot(1:20, unlist(m[,2]), lwd=2, type="o", col='blue', ylab="MSE")

```
