---
output:
  html_document: default
  pdf_document: default
---
# Clustering
### Karen Mazidi

Modified from Kabacoff, "R in Action", 2nd ed

## K-means

Apply k-means to the wine data set, which contains 13 chemical measurements on 178 Italian wines. 

The first column, Type, indicates 1 or 3 wine varieties. We will drop this variable for the clustering.

```{r}
data(wine, package="rattle")
names(wine)
head(wine)
df <- scale(wine[-1])
head(df)
```

Write a function to plot the within-groups sums of squares vs. the number of clusters. 


```{r}
wsplot <- function(data, nc=15, seed=1234){
  wss <- (nrow(data)-1)*sum(apply(data,2,var))
  for (i in 2:nc){
    set.seed(seed)
    wss[i] <- sum(kmeans(data,centers=i)$withinss)
  }
  plot(1:nc, wss, type="b", xlab="Number of Clusters",
       ylab="Within groups sum of squares")
}
wsplot(df)
```


Use the NbClust() function to help determine the best number of clusters.

In the within-groups plot, we see an "elbow" around 3, suggesting that 3 clusters is a good choice. 

```{r}
library(NbClust)
set.seed(1234)
nc <- NbClust(df, min.nc=2, max.nc=15, method="kmeans")
table(nc$Best.n[1,])
barplot(table(nc$Best.n[1,]),
        xlab="Number of Clusters", ylab="Number of Criteria",
        main="Number of Clusters Chosen by 26 Criteria")
```

## KMeans

Fit the model using the kmeans() function. We set a seed first so we get reproducible results. 

The centroids are found in fit.km$centers and we display those. 

```{r}
set.seed(1234)
fit.km <- kmeans(df, 3, nstart=25)
fit.km$size
fit.km$centers

```

The centroids were calculated based on the scaled data. Next we use the aggregate() function along with the cluster membership to get variable means for each cluster in units of the original, unscaled, data. 


```{r}
aggregate(wine[-1], by=list(cluster=fit.km$cluster), mean)

```

## Model Analysis

If we cross-tabulate the Type in column 1 of the wine data with cluster membership, we see that the clusters are strongly correlated with the wine type. 


```{r}
ct.km <- table(wine$Type, fit.km$cluster)
ct.km
```

We can quantify the agreement between the type and the cluster using an adjusted Rand index. The adjusted Rand index provides a measure of the agreement between two partitions, adjusted for chance. The range of the index is from -1 (no agreement) to +1 (perfect agreement). 

The results below show very good agreement!

```{r}
library(flexclust)
randIndex(ct.km)

```

