---
title: "Introduction to ggplot2"
author: "Karen Mazidi"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

There are 7 grammatical elements in ggplot2, the first 3 of these are essential to getting something plotted:


* data - the data being plotted should be the first argument, or specify data=...
* aesthetics - the scales onto which we plot; use aes() to specify at least x= and y= if needed as well as other parameters for customization
* geometries - visual elements such as points, lines, etc.
* facets - for plotting multiples
* statistics - representations to aid understanding
* coordinates - space on which data will be plotted
* themes - you can customize your own theme to use over and over

### load tidyverse and some data

Loading the diabetes data set from package mlbench.

```{r, message=FALSE}
library(tidyverse)
library(mlbench)
data("PimaIndiansDiabetes2")

tb <- tbl_df(PimaIndiansDiabetes2)
```


## Explore ggplot2

Hadley Wickham developed ggplot2 in 2005, inspired by a grammar of graphics developed by Leland Wildinson in 1999. The ggplot2 functions are much more powerful than standard R graphs but also slower. 

We have a short example below showing important components of building a ggplot. First we specify the data, then the aesthetics which are how the data is represented, followed by the geometry and finally labels.

```{r, message=FALSE}
ggplot(tb, aes(x=mass, y=glucose)) +
  geom_point() +
  labs(title="Glucose and BMI", x="BMI", y="Glucose")
```
Next we add some color and a smoothing line which helps us see a trend in the data. By default the smoothing line to highlight the trend in the data

```{r}
ggplot(tb, aes(x=mass, y=glucose)) +
  geom_point(pch=20, color='blue', size=1.5) +
  geom_smooth(method='lm', color='red', linetype=2) +
  labs(title="Glucose and BMI", x="BMI", y="Glucose")
```

### informative graph


```{r}
ggplot(tb, 
  aes(x=tb$mass, y=tb$age, shape=diabetes, col=pregnant)) +
  geom_point(size=2) +
  labs(x="BMI",    y="Age")
```



### facet_grid

* filter out rows with NAs in glucose or insulin
* create 2 new factor columns, glucose_high and insulin_high
* plot

The facet grid for 2 binary variables has 4 windows for all combinations. 

```{r}
tb <- filter(tb, !is.na(glucose), !is.na(insulin))

tb <- mutate(tb, glucose_high = factor(ifelse(tb$glucose>mean(tb$glucose), 1, 0)))
tb <- mutate(tb, insulin_high = factor(ifelse(tb$insulin>mean(tb$insulin), 1, 0)))

ggplot(tb, 
  aes(x=mass, y=age, shape=diabetes, col=pregnant)) +
  geom_point(size=2) +
 facet_grid(glucose_high~insulin_high)       
```


### histogram

```{r}
ggplot(tb, aes(x=mass)) +
  geom_histogram(fill="cornsilk4")
```

### boxplot and rug

```{r}
ggplot(tb, aes(x=diabetes, y=mass)) +
  geom_boxplot(notch=TRUE) +
  geom_point(position="jitter", color="cornflowerblue", alpha=.5) +
  geom_rug(color="cornflowerblue")
```

### density plot

```{r}
ggplot(tb, aes(x=mass, fill=diabetes)) +
  geom_density(alpha=0.5)
```
### bubble chart

```{r}
ggplot(tb, 
       aes(x=mass, y=glucose, size=pregnant)) +
  geom_point(shape=21, fill="cornflowerblue")
```

### grid

```{r}
library(gridExtra)
p1 <- ggplot(tb, aes(x=insulin_high)) + geom_bar(fill="cornflowerblue") 
p2 <- ggplot(tb, aes(x=glucose_high)) + geom_bar(fill="cornflowerblue")
grid.arrange(p1, p2, ncol=2)
```

